Ideas cascading into keystrokes -- modelling writing hesitations as finite mixture process

### 2024-05-02 report/manuscript.Rmd
update tutorial final for now

csl               : "apa.csl"

author: 
  - name          : "Jens Roeser"
    affiliation   : "1"
    address       : "50 Shakespeare St, Nottingham NG1 4FQ"
    corresponding : yes 
    email         : "jens.roeser@ntu.ac.uk"
  - name          : "Mark Torrance"
    affiliation   : "1"
  - name          : "Rianne Conijn"
    affiliation   : "2"
  - name          : "Evgeny Chukharev"
    affiliation   : "3"

affiliation:
  - id            : "1"
    institution   : "Department of Psychology, Nottingham Trent University, United Kingdom"

  - id            : "2"
    institution   : "Artificial Intelligence Systems Institute, Eindhoven University of Technology, The Netherlands"

  - id            : "3"
    institution   : "Department of English, Iowa State University, Iowa"

@@

keywords: "Cascading parallel models; serial models; finite mixture models; Bayesian models; text composition; keystrokes"

@@

documentclass     : "apa7"
classoption       : "man"
output            : 
  papaja::apa6_pdf:
    keep_tex: TRUE

@@

header-includes:
  - \usepackage{booktabs}
  - \usepackage{longtable}
  - \usepackage{graphicx}
  - \usepackage{array}
  - \usepackage{multirow}
  - \usepackage{float}
  - \usepackage{colortbl}
  - \usepackage{threeparttable}
  - \usepackage[normalem]{ulem}
  - \usepackage[utf8]{inputenc}
  - \usepackage{icomma}
  - \usepackage{pdflscape}
  - \newcommand{\blandscape}{\begin{landscape}}
  - \newcommand{\elandscape}{\end{landscape}}
  - \DeclareCaptionFormat{cont}{#1 (cont.)#2#3\par}
---

@@

```{=html}
<!-- TODO: 
- address difference between posteriors of shift analysis for constrained and unconstrained model?
- pausing frequency for words and sentences are equal but more importantly pauses occure less than 80% of the time indicating parallel planing.
- would be really good, on some point, to distinguish between open and closed class words
-->
```

@@

```{=html}
<!--

----
Add to OSF
Work on tutorial and add as link to methods
----

~~~~~
Say how the results don't contradic the reported studies but speaks better to their conclusions 
but the modelling frameword doesn't require us to impose pause thresholds
~~~~~

~~~~
Move from results to discussion?
Open: to what extent are even short latencies longer at sentences. This is obvious for when multiple keystrokes are summed across at sentences boundaries as there are typically three key transitions necessary to start a sentence: (1) .^_ , (2) _^[shift], (3) [shift]^C
We believe the second one is the linguistically meaningful one
I think give Davids papers as examples, if he uses sums instead of means (other studies?)

~~~

@@

~~~~
I think the serial account isn't all inconsistent with the mixture model. I don't think the serial account says that pauses must occure on linguistic edges but between bursts which can in principlpe happen anywhere (but why would they be random). Obviously the problem here is that one needs to postulate what a pause is instead of leaving it to the model to work out. The serial account and the cascading account make very similar predictions for inexperienced writers.

Maybe distinguih between a strong version of serial model and a weak version that is basically a chunky cacading model

-->
```

@@

\setlength{\parindent}{-0.5in}

\setlength{\leftskip}{0.5in}

<div id = "references"></div>

>>>>

csl               : "`r system.file('rmd', 'apa7.csl', package = 'papaja')`"
documentclass     : "apa7"
output            : 
  papaja::apa6_pdf:
    keep_tex: FALSE

author: 
  - name          : "Jens Roeser"
    affiliation   : "1"
    address       : "50 Shakespeare St, Nottingham NG1 4FQ"
    corresponding : yes 
    email         : "jens.roeser@ntu.ac.uk"
  - name          : "Mark Torrance"
    affiliation   : "1"
  - name          : "Rianne Conijn"
    affiliation   : "2"
  - name          : "Gunn Helen Ofstad"
    affiliation   : "3"
  - name          : "Evgeny Chukharev"
    affiliation   : "4"

affiliation:
  - id            : "1"
    institution   : "Department of Psychology, Nottingham Trent University, United Kingdom"
  - id            : "2"
    institution   : "Artificial Intelligence Systems Institute, Eindhoven University of Technology, The Netherlands"
  - id            : "3"
    institution   : "Educational Sciences and Humanities, University of Stavanger, Norway"
  - id            : "4"
    institution   : "Department of English, Iowa State University, Iowa"

@@

keywords: "Cascading semi-parallel models; serial models; finite mixture models; Bayesian models; text composition; keystrokes"

@@

@@

header-includes:
  - \usepackage{booktabs}
  - \usepackage{longtable}
  - \usepackage{graphicx}
  - \usepackage{array}
  - \usepackage{multirow}
  - \usepackage{float}
  - \usepackage{colortbl}
  - \usepackage{threeparttable}
  - \usepackage[normalem]{ulem}
  - \usepackage[utf8]{inputenc}
  - \usepackage{icomma}
  - \usepackage{pdflscape}
  - \newcommand{\blandscape}{\begin{landscape}}
  - \newcommand{\elandscape}{\end{landscape}}
  - \DeclareCaptionFormat{cont}{#1 (cont.)#2#3\par}
  - |
    \makeatletter
    \renewcommand{\paragraph}{\@startsection{paragraph}{4}{\parindent}%
      {0\baselineskip \@plus 0.2ex \@minus 0.2ex}%
      {-1em}%
      {\normalfont\normalsize\bfseries\typesectitle}}

    \renewcommand{\subparagraph}[1]{\@startsection{subparagraph}{5}{1em}%
      {0\baselineskip \@plus 0.2ex \@minus 0.2ex}%
      {-\z@\relax}%
      {\normalfont\normalsize\bfseries\itshape\hspace{\parindent}{#1}\textit{\addperi}}{\relax}}
    \makeatother

---

@@

<!-- TODO: 
- would be really good, on some point, to distinguish between open and closed class words
-->

@@

<!--

Add to OSF
Work on tutorial and add as link to methods
Say how the results don't contradic the reported studies but speaks better to their conclusions

@@

-->

@@

<!-- \setlength{\parindent}{-0.5in} -->

<!-- \setlength{\leftskip}{0.5in} -->

<div id = "refs"></div>
